---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cosr

<!-- badges: start -->
<!-- badges: end -->

The goal of `cosr` is to facilitate the exploration and analysis of open science practices and resource lifecycles using data from the Open Science Framework.

Unlike [`osf.io`](https://github.com/CenterForOpenScience/osf.io/tree/develop/osf), `cosr` is:

- An R package, not a web application/service.
- Built for data analysis and exploration, not database management and software development.
- Tailored for researchers and data scientists interested in open science metrics.
- Provides functions to process and analyze OSF data programmatically, reproducibly, and interactively.
- Designed to work with columnar data formats (e.g., Parquet) for efficient handling of large datasets.

## Installation

You can install the development version of `cosr` like so:

``` r
#install.packages("devtools")
devtools::install_github("alexjakubow/cosr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(cosr)
## basic example code
```