---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cosr

<!-- badges: start -->
<!-- badges: end -->

The goal of `cosr` is to facilitate the exploration and analysis of open science practices and resource lifecycles using data from the Open Science Framework.

Unlike [`osf.io`](https://github.com/CenterForOpenScience/osf.io/tree/develop/osf), `cosr` is:

- An R package, not a web application/service.
- Built for data analysis and exploration, not database management and software development.
- Tailored for researchers and data scientists interested in open science metrics.
- Provides functions to process and analyze OSF data programmatically, reproducibly, and interactively.
- Designed to work with columnar data formats (e.g., Parquet) for efficient handling of large datasets.

## Installation

You can install the development version of `cosr` like so:

``` r
#install.packages("devtools")
devtools::install_github("alexjakubow/cosr")
```

## Package Architecture

The `cosr` package is organized by user experience level to make it easy to find the functions you need:

- **workflows.r** 🟢 - High-level pipeline functions (start here!)
- **queries.r** 🟡 - Individual attribute extraction functions
- **analysis.r** 🟡 - Analytical and summarization functions
- **reporting.r** 🟡 - Reporting dataset generation and Google Sheets export
- **compute.r** 🔴 - Low-level computational logic (advanced)
- **utils.r** 🔴 - Infrastructure and utilities (advanced)
- **globals.r** - Global variable declarations (for R CMD check)

### Typical Workflow

The complete workflow consists of 5 steps:

```r
library(cosr)

# Step 1: Generate current attribute memberships for all registrations
assign_attributes(sample = expr_valid_regs)

# Step 2: Generate monthly LOS time series
event_history_monthly(
  .start = "2022-09-01",
  .end = Sys.Date(),
  .fileout = "data/los_ts.parquet"
)

# Step 3: Summarize by all attributes (including aggregate overall summary)
los_summary_by_attributes_all(
  subgroup_vars = c(SUPPORTED_ATTRIBUTES, "overall"),
  los_ts_path = "data/los_ts.parquet"
)
# Result: data/los_ts_overall.parquet + data/los_ts_<attribute>.parquet

# Step 4: Generate tidy reporting dataset
report_df <- generate_reporting_dataset()
arrow::write_parquet(report_df, "data/reporting_dataset.parquet")

# Step 5: Publish to Google Sheets (optional)
googlesheets4::gs4_auth()
annual <- format_gsheet_annual(report_df)
write_gsheet_report(annual, sheet_name = "LOS Indicators on OSF")
```

See `vignette("generate-datasets")` for data generation and `vignette("reporting-dataset")` for reporting workflows.

### Available Attributes

The package supports 8 registration attributes for subgroup analysis:

- `institution` - OSF Institution (OSFI) affiliations
- `affiliated` - Binary OSFI affiliation status
- `funder` - Funding organization names
- `funded` - Binary funded/unfunded status
- `subject` - Subject/Discipline classifications
- `subject_parent` - Parent subject categories
- `provider` - OSF Provider (i.e., registry)
- `template` - Registration template/schema

## Example

Here's a quick example of analyzing LOS status by institution:

```{r example, eval=FALSE}
library(cosr)
library(dplyr)

# Get current institutional affiliations
institutions <- get_registration_institution(lazy = FALSE)

# Get current LOS status
los_status <- resource_counts_at_date(
  sample = expr_valid_regs, # see R/utils.r for sample parameters
  cutoff = Sys.Date(),
  lazy = FALSE
)

# Combine and summarize
institutions |>
  left_join(los_status, by = "node_id") |>
  group_by(institution) |>
  summarise(
    n_regs = n(),
    n_los = sum(is_los),
    prop_los = n_los / n_regs
  ) |>
  arrange(desc(prop_los))
```