---
title: "Trends in Lifecycle Open Science on OSF"
date: "2026-02-11"
date-modified: last-modified
format:
  html:
    output-file: trends-html.html
    embed-resources: true
    toc: true
    toc-depth: 3
    theme: default
    echo: false
    warning: false
    message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE # Set to TRUE when data files are available
)
```

# Overview

This document demonstrates how to analyze trends in Lifecycle Open Science (LOS) on OSF using data generated by the `cosr` package workflows.

**Note:** This analysis requires that you first run the data generation workflow. The code examples below assume you have already generated the LOS time series data files.

# Prerequisites

Before running this document, you need to:

1. Run `assign_attributes()` to generate attribute files
2. Run `event_history_monthly()` to generate time series data
3. Run `los_summary_by_attributes_all()` to generate summary data

# Setup

```{r libraries, eval=TRUE}
library(dplyr)
library(lubridate)
library(timetk)
library(arrow)
library(cosr)
library(gt)
```

# Helper Functions

```{r functions, eval=TRUE}
# Function to create summary tables by attribute
table_one <- function(var = NULL, data_dir = "data") {
  # Check if data directory exists
  if (!dir.exists(data_dir)) {
    stop(
      "Data directory '",
      data_dir,
      "' not found. ",
      "Please run the data generation workflow first.",
      call. = FALSE
    )
  }

  # Source the data
  if (is.null(var)) {
    var = "Overall"

    # Check if file exists
    file_path <- file.path(data_dir, "los_ts_overall.parquet")
    if (!file.exists(file_path)) {
      stop(
        "File '",
        file_path,
        "' not found. ",
        "Please run event_history_monthly() first.",
        call. = FALSE
      )
    }

    tbl <- arrow::open_dataset(file_path) |>
      arrow::to_duckdb() |>
      dplyr::filter(date == max(date, na.rm = TRUE)) |>
      dplyr::select(
        n_osr,
        are_los,
        prop_los,
        have_outputs,
        prop_outputs,
        have_outcomes,
        prop_outcomes
      )
  } else {
    # Check if file exists
    file_path <- file.path(
      data_dir,
      paste0("los_ts_", tolower(var), ".parquet")
    )
    if (!file.exists(file_path)) {
      stop(
        "File '",
        file_path,
        "' not found. ",
        "Please run los_summary_by_attributes_all() first.",
        call. = FALSE
      )
    }

    tbl <- arrow::open_dataset(file_path) |>
      arrow::to_duckdb() |>
      dplyr::filter(date == max(date, na.rm = TRUE)) |>
      dplyr::select(
        .data[[var]],
        n_osr,
        are_los,
        prop_los,
        have_outputs,
        prop_outputs,
        have_outcomes,
        prop_outcomes
      )
  }

  # Format the table
  tbl |>
    dplyr::collect() |>
    gt::gt() |>
    gt::fmt_percent(
      columns = starts_with("prop_"),
      decimals = 1
    ) |>
    gt::fmt_number(
      columns = starts_with("n_"),
      decimals = 0,
      use_seps = TRUE
    ) |>
    gt::cols_label(
      n_osr = "Total",
      have_outputs = "w/ Outputs",
      have_outcomes = "w/ Outcomes",
      are_los = "LOS",
      prop_outputs = "% w/ Outputs",
      prop_outcomes = "% w/ Outcomes",
      prop_los = "% LOS"
    ) |>
    gt::tab_header(
      title = glue::glue(
        "Lifecycle Open Registrations on OSF by {stringr::str_to_title(var)}"
      )
    ) |>
    gt::opt_interactive(
      use_sorting = TRUE,
      use_search = TRUE
    )
}
```

# Lifecycle Open Science Today on OSF

## Overall Summary

```{r overall-table}
# Note: This will only work if you've run the data generation workflow
table_one()
```


## Summary by Attribute


### By Affiliation Status

```{r affiliated-table}
table_one("affiliated")
```

### By Institution

```{r institution-table}
table_one("institution")
```

### By Funder

```{r funder-table}
table_one("funder")
```

### By Funded Status

```{r funded-table}
table_one("funded")
```

### By Registration Template

```{r template-table}
table_one("template")
```

### By Provider

```{r provider-table}
table_one("provider")
```

### By Subject

```{r subject-table}
table_one("subject")
```

### By Parent Subject

```{r subject-parent-table}
table_one("subject_parent")
```

# Trends in Lifecycle Open Science on OSF

## Time Series Visualization

```{r ts-plot-setup}
# Load time series data (after running los_summary_by_attributes_all())
data_dir <- "data"

if (file.exists(file.path(data_dir, "los_ts_affiliated.parquet"))) {
  osfi_ts <- arrow::open_dataset(
    file.path(data_dir, "los_ts_affiliated.parquet")
  ) |>
    collect()
} else {
  message(
    "Data file not found. Please run los_summary_by_attributes_all() first."
  )
}
```

### LOS Counts by Affiliation

```{r ts-counts-faceted}
# Faceted time series plot of LOS counts by affiliation
osfi_ts |>
  plot_time_series(
    as.POSIXct(date),
    are_los,
    .color_var = year(as.POSIXct(date)),
    .facet_vars = affiliated,
    .smooth = FALSE,
    .interactive = TRUE,
    .title = "Lifecycle Open Registrations by Affiliation",
    .y_lab = "Count",
    .x_lab = "Date"
  )
```

### LOS Proportions by Affiliation

```{r ts-proportions-comparison}
# Comparison time series plot of LOS proportions by affiliation
osfi_ts |>
  plot_time_series(
    as.POSIXct(date),
    prop_los,
    .color_var = affiliated,
    .smooth = FALSE,
    .interactive = TRUE,
    .title = "Proportion of OSRs that are Lifecycle Open by Affiliation",
    .y_lab = "LOS/Total",
    .x_lab = "Date"
  )
```
