---
title: "Lifecycle Open Science on OSF"
date: "2026-02-05"
date-modified: last-modified
---

# Overview

This method measures lifecycle open science on OSF by looking at open science practices linked to open science (i.e., "valid") registrations.  

## Definitions

> **Open Science Registration (OSR)**: An OSF Registration that is jointly *open*, *non-deprecated* and *authentic*

- Openness Criteria:
    1. Public (open-1)

- Non-Deprecation Criteria:
    1. Registered (nondeprecated-1)
    2. Accepted (nondeprecated-2)
    3. Not Deleted (nondeprecated-3)

- Authenticity Criteria:
    1. Not Spam (flagged or confirmed) (authentic-1)


> **LOS-Reg**: An OSR that represents a lifecycle open science research project.

- Lifecycle Opennness Criteria:

    1. **Output** - The OSR has at least 1 linked [Open Practice Resource](https://help.osf.io/article/452-open-practice-badges) from among: "Data", "Analytic Code", "Materials" or "Supplements".  
    2. **Outcome** - The OSR has at least 1 linked "Papers" Open Practice Resource.


# Summary Profiles of Current OSRs

```{r}
library(cosr)
library(dplyr)
library(purrr)

# Get Current Population of OSRs
osr <- open_parquet(tbl = "osf_abstractnode") |>
  filter(type == "osf.registration", !!!expr_valid_regs)
OSF_TOTAL <- osr |>
  summarise(n = n()) |>
  pull(n)

# Compute Subpopulations
osr_subpops <- registration_summaries()
```

As of `r Sys.Date()`, there are `r OSF_TOTAL` Open Science Registrations (OSRs) on OSF. The following sections present descriptive profiles of these OSRs by different subpopulations (segments).


## Institutional Affiliation
OSFI STATUS
```{r}
osr_subpops$`Affiliation Status`
```


INSTITUTION
```{r}
osr_subpops$Institution
```


## Registry and Template

REGISTRY
```{r}
osr_subpops$`Registry`
```


TEMPLATE
```{r}
osr_subpops$`Registration Template`
```


## Subject and Funder

SUBJECT
```{r}
osr_subpops$`Subjects`
```

SUBJECT DETAIL
```{r}
osr_subpops$`Subject Detail`
```

FUNDER STATUS
```{r}
osr_subpops$`Funder Status`
```


FUNDER DETAIL
```{r}
osr_subpops$`Funder`
```