---
title: "Creating Lifecycle Open Science Time Series Data"
date: "2026-02-11"
date-modified: last-modified
vignette: >
  %\VignetteIndexEntry{Creating LOS Time Series Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

# Overview

This vignette demonstrates the typical workflow for creating Lifecycle Open Science (LOS) time series datasets using the `cosr` package.

The package is organized by user experience level:

- **workflows.r** - High-level pipeline functions (start here!)
- **queries.r** - Individual attribute extraction functions
- **analysis.r** - Analytical and summarization functions
- **compute.r** - Low-level computational logic
- **utils.r** - Infrastructure and utilities

# Typical Workflow

The standard workflow consists of three main steps using high-level functions from `workflows.r`:

## Step 1: Generate Current Attribute Memberships

First, we extract the current attributes (subgroup memberships) for all valid Open Science Registrations (OSRs). This includes attributes like institution, funder, template, provider, etc.

```{r eval=FALSE}
library(cosr)

# Generate all 8 supported attributes and cache to data/
assign_attributes(sample = expr_valid_regs)
```

This creates Parquet files in `data/`:
- `osr_affiliated.parquet`
- `osr_funder.parquet`
- `osr_funded.parquet`
- `osr_institution.parquet`
- `osr_template.parquet`
- `osr_provider.parquet`
- `osr_subject.parquet`
- `osr_subject_parent.parquet`

## Step 2: Generate Monthly LOS Time Series

Next, we create a monthly time series tracking resource linkages and LOS status over time.

```{r eval=FALSE}
# Generate monthly snapshots from Jan 2023 to present
event_history_monthly(
  .start = "2023-01-01",
  .end = Sys.Date(),
  .sample = expr_valid_regs,
  .fileout = "data/los_ts.parquet"
)
```

This creates `data/los_ts.parquet` containing monthly snapshots with:
- Resource counts (linked/removed outputs and outcomes)
- LOS status indicators
- LOS change tracking (gains, losses, no change)

## Step 3: Summarize by All Subgroups

Finally, we compute summary statistics for each attribute over time.

```{r eval=FALSE}
# Generate summaries for all supported attributes
los_summary_by_attributes_all(
  subgroup_vars = SUPPORTED_ATTRIBUTES,
  los_ts_path = "data/los_ts.parquet"
)
```

This creates time series summary files:
- `data/los_ts_affiliated.parquet`
- `data/los_ts_funder.parquet`
- `data/los_ts_funded.parquet`
- `data/los_ts_institution.parquet`
- `data/los_ts_template.parquet`
- `data/los_ts_provider.parquet`
- `data/los_ts_subject.parquet`
- `data/los_ts_subject_parent.parquet`

# Advanced Usage

## Custom Attribute Analysis

For analyzing a single attribute with custom settings:

```{r eval=FALSE}
# Summarize by institution only
institution_summary <- los_summary_by_attribute(
  subgroup_var = "institution",
  los_ts_path = "data/los_ts.parquet"
)
```

## Point-in-Time Snapshots

To get LOS status at a specific date:

```{r eval=FALSE}
# Get LOS status as of January 1, 2024
snapshot_jan2024 <- resource_counts_at_date(
  sample = expr_valid_regs,
  cutoff = "2024-01-01",
  lazy = FALSE
)
```

## Custom Filtering

To analyze specific subgroups:

```{r eval=FALSE}
# Get only funded registrations
funded_only <- get_registration_funded(lazy = FALSE) |>
  dplyr::filter(funded == "Funded")
```

# Output Files Summary

After running the complete workflow, you'll have:

**Attribute files (8):**
- Current subgroup memberships for all registrations

**Time series files (9):**
- Monthly LOS time series (main file)
- Monthly summaries by each of 8 attributes

**Metadata (1):**
- Funder metadata lookup table

These files can then be used for downstream visualization and analysis.