---
title: "Generating Summary Datasets for Reporting"
date: "2026-02-18"
date-modified: last-modified
vignette: >
  %\VignetteIndexEntry{Generating Summary Datasets for Reporting}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

This vignette demonstrates how to generate tidy, long-form summary datasets
suitable for reporting and visualization, and how to publish those datasets to
Google Sheets in a format that facilitates spreadsheet-based analysis and inference.

**Note:** This vignette requires that you have already run the data generation
workflow described in `vignette("generate-datasets")`.

# Prerequisites

The following files must exist in your `data/` directory before running Steps
4 and 5:

- `data/los_ts_summary.parquet` — aggregate time series (Step 3)
- `data/los_ts_<attribute>.parquet` — one file per attribute (Step 3)

See `vignette("generate-datasets")` for how to generate them.

# Setup

```{r libraries, eval=TRUE}
library(cosr)
library(dplyr)
library(arrow)
```

---

# Step 4 — Generate the Reporting Dataset

[`generate_reporting_dataset()`] combines the aggregate overall summary with
per-attribute subgroup summaries into a single tidy long-form data frame. This
is the canonical dataset for downstream analysis and visualization.

```{r step-4}
report_df <- generate_reporting_dataset()

# Save to disk
arrow::write_parquet(report_df, "data/reporting_dataset.parquet")
```

## Dataset Structure

| Column | Description | Example values |
|---|---|---|
| `attribute` | Subgroup dimension | `"overall"`, `"institution"`, `"funded"` |
| `subgroup` | Subgroup value | `"MIT"`, `"Funded"`, `NA` (for overall) |
| `date` | Data date | `"2024-01-01"` |
| `metric` | Metric name (prefix stripped) | `"osr"`, `"outputs"`, `"los"` |
| `moment` | Type of value | `"value"`, `"lag"`, `"delta"`, `"growth"` |
| `value` | Numeric value | `1234`, `0.042` |
| `value_type` | Count or proportion | `"count"`, `"proportion"` |

## Moments

The `moment` column distinguishes four temporal perspectives on each metric:

- **`value`** — the observed metric value at the given date
- **`lag`** — the value from the prior period (month)
- **`delta`** — the absolute change from the prior period (`value - lag`)
- **`growth`** — the relative change from the prior period (`delta / lag`)

## Value Types

The `value_type` column normalizes metric name prefixes so that counts and
proportions for the same concept share a common `metric` value:

| Original metric | `metric` | `value_type` |
|---|---|---|
| `n_osr` | `osr` | `count` |
| `are_los` | `los` | `count` |
| `have_outputs` | `outputs` | `count` |
| `have_outcomes` | `outcomes` | `count` |
| `prop_los` | `los` | `proportion` |
| `prop_outputs` | `outputs` | `proportion` |
| `prop_outcomes` | `outcomes` | `proportion` |

## Common Queries

```{r queries}
# Overall current snapshot (latest date)
report_df |>
  filter(attribute == "overall", moment == "value", date == max(date))

# LOS proportion by institution at latest date
report_df |>
  filter(
    attribute == "institution",
    metric == "los",
    value_type == "proportion",
    moment == "value",
    date == max(date)
  ) |>
  arrange(desc(value))

# Monthly LOS growth rates overall
report_df |>
  filter(
    attribute == "overall",
    metric == "los",
    value_type == "count",
    moment == "growth"
  ) |>
  arrange(date)
```

---

# Step 5 — Publish to Google Sheets

Step 5 formats the overall (aggregate) reporting data into a wide annual layout
matching the "LOS Indicators on OSF" tracking spreadsheet, and writes each year
to its own tab in a Google Sheet.

## Sheet Format

Each tab corresponds to one calendar year and contains three subtables stacked
vertically, separated by two blank rows:

```
TOTALS        |  1  |  2  |  3  | ... | 12
─────────────────────────────────────────────
Total Registrations
Registrations with Linked Outputs
Registrations with Linked Outcomes
LOS Registrations (Regs with Outputs and Outcomes)
% of Registrations that are LOS
[blank]
[blank]
NET CHANGE    |  1  |  2  |  3  | ... | 12
─────────────────────────────────────────────
...
[blank]
[blank]
% GROWTH      |  1  |  2  |  3  | ... | 12
─────────────────────────────────────────────
...
```

Month columns are blank for months that have not yet occurred.

## Authentication

`googlesheets4` uses OAuth 2.0. Authenticate once per session before writing:

```{r auth}
googlesheets4::gs4_auth()
```

## Format and Write

```{r step-5}
# Format: produces a named list, one data frame per year
annual <- format_gsheet_annual(report_df)

# Inspect a single year before writing
annual[["2025"]]

# Write all years to a new Google Sheet (one tab per year)
write_gsheet_report(
  annual_list = annual,
  sheet_name = "LOS Indicators on OSF"
)
```

## Update an Existing Sheet

If the sheet already exists, supply its ID or URL to overwrite tabs in place
rather than creating a new sheet:

```{r step-5-update}
write_gsheet_report(
  annual_list = annual,
  sheet_id = "your-google-sheet-id-here"
)
```

The Sheet ID is the long alphanumeric string in the Google Sheets URL:
`https://docs.google.com/spreadsheets/d/**<sheet-id>**/edit`

## Write a Single Year

```{r step-5-single-year}
# Format only 2025 data
annual_2025 <- format_gsheet_annual(report_df, years = 2025)

write_gsheet_report(
  annual_list = annual_2025,
  sheet_id = "your-google-sheet-id-here"
)
```

---

# Complete Workflow (Steps 1–5)

```r
library(cosr)

# Step 1: Generate attribute files
assign_attributes(sample = expr_valid_regs)

# Step 2: Generate monthly time series
event_history_monthly(
  .start   = "2023-01-01",
  .end     = Sys.Date(),
  .fileout = "data/los_ts.parquet"
)

# Step 3: Summarize by all attributes and in aggregate
los_summary_by_attributes_all(
  los_ts_path = "data/los_ts.parquet"
)

# Step 4: Generate reporting dataset
report_df <- generate_reporting_dataset()
arrow::write_parquet(report_df, "data/reporting_dataset.parquet")

# Step 5: Publish to Google Sheets
googlesheets4::gs4_auth()
annual <- format_gsheet_annual(report_df)
write_gsheet_report(annual, sheet_name = "LOS Indicators on OSF")
```
